### 2018-10-27 Paper/revision/reviewer_replies.tex

Comments and Questions:
\begin{enumerate}

\item \referee{P16, Figure.2,  the estimated screening parameters eq (5) based on eq(1)
quickly converge with a small number of samples (5-10 samples). It seems that
if $\mu_i$ is based on eq(1), it will still require a reasonable amount of
samples, can authors comment on why this example converges with a small number
of samples? The same question applies to Figure 5 and Figure 7.}

Convergence behavior in this case depends upon the variability in the gradient of
$G(\theta)$ with respect to individual parameters, $\theta_i$ in their respective
domains. For the numerical tests and the application considered in this work, the
gradient is not observed to vary significantly in the considered domain of the
parameters. In other words, the variance of the estimator is low. 
Hence, a small number of Monte Carlo samples are able to estimate
the screening parameters. This is a common observation in many practical
applications, and therefore the proposed methodology involving an iterative
approach can yield significant computational savings.

These points have been added in the discussion in Section~6 (Summary and
Conclusion, Page 34) in the revised manuscript.

@@

The reviewer correctly points out that the resulting surrogate model is either an RSS or
an FSS. However, we would like to clarify that the main purpose of the proposed
methodology is to construct an RSS based on estimates of a screening metric that
involves derivative-based global sensitivity measure (DGSM) for each uncertain parameter.
However, since estimating the metric requires model evaluations, we further exploit these
evaluations to construct an FSS (free of cost) using a sparse regression-based approach
as mentioned in the paper. Note that the FSS construction is performed simultaneously to 
ensure that the computational effort associated with obtaining converged estimates of
the screening metric and RSS construction does not overshoot that required for
constructing a sufficiently accurate FSS.

The above points are already included in the paragraph on the discussion of the proposed
methodology in Section 3. Nevertheless, based on the reviewer's feedback, we have tried to
clarify and emphasize these points in the first paragraph in Section 6 (Summary and Conclusion)
in the revised manuscript.

\item
\referee{The reviewer would also like to see the comparison of the computational cost
(not just the numerical results) of DGSMs to variance based sensitivity indices
with increasing parameter dimension, e.g. Sobol’ indices incorporated with
sparse grid, since the difference decreases with increasing number of parameters.}

As suggested by the reviewer, the computational cost pertaining to the estimation of DGSM-based
sensitivity indices using the approach presented in this work is compared with the cost associated 
with estimating the total-effect Sobol' indices using a sparse-basis PCE
for the 19-dimensional reaction kinetics application. Additionally, the comparison is 
performed for the same application with 33 uncertain rate-controlling parameters to investigate
the impact of dimensionality.

For the 19-dimensional case, the DGSM-based approach seems 
slightly more efficient depending upon the model runs required for training and verifying the sparse-basis PCE
used to compute the total-effect Sobol' indices. However, computational gains are observed to 
be much more substantial in the case of DGSM-based approach for the 33-dimensional problem
owing to the increase in the number of training points required to construct the PCE. Our analysis
clearly indicates that with increasing dimensionality, the DGSM-based approach presented in this
work is expected to yield a higher computational dividend especially when efficient gradient estimation
techniques are employed.

A new sub-section (5.3) on ``comparative cost analysis" comprising the results and
discussion based on the above points has been included in the revised
manuscript.

@@

\item 
\referee{In section 4 and 5, no predictive convergence using the test data sample is
presented, only the convergence of $\epsilon_\text{LOO}$, which reuses the training data as
the validation set, is shown. This shows only the convergence of the model,
and does not necessarily indicate the (predictive) accuracy of the constructed
model, hence could lead to false conclusion. The reviewer would like to see the
convergence of $\epsilon_{L_2}$.}

As suggested by the reviewer, the plots and related discussion pertaining to the convergence 
of $\epsilon_{L_2}$ have been included for the numerical tests presented in Section 3 (Figures 3b
and 6b) and 
the chemical kinetics application presented in Section 4 (Figure 8, right). As expected, the predictive
accuracy of the surrogates is observed to increase with the number of training points in all cases. 
Additionally, the RSS is observed to be more accurate when a small number of training samples, $N$=10
is used. However, as $N$ increases the FSS is observed to be more accurate depending upon the
application.

@@

\item 
\referee{How would the proposed approach behave when presented with severe nonlinearity
or rare events where the screening metrics could misrepresent the significance
of parameters?}

Severe non-linearity in the model output as a function of uncertain inputs could potentially lead
to a significantly large variability in its gradient. Hence, a large number of Monte Carlo samples might be
required to estimate the screening metrics with reasonable accuracy in that case depending upon the
application. Therefore, the computational effort is expected to increase in such cases. In the presence
of rare events, the variance-based approaches typically lead to inaccurate estimates of 
parametric sensitivities. Since the proposed methodology in this work aims to reduce the computational
effort pertaining to total-effect Sobol' index which is a variance-based measure, it might lead to
inaccurate estimates of the screening metric and therefore to erroneous relative importance of the uncertain
inputs.

### 2018-10-28 Paper/app.tex
..

The computational cost is estimated in terms of the number of function evaluations or model runs,
denoted by $\mathcal{M}$ in each case. In the case of DGSM-based approach presented in this work, 
$\mathcal{M}$ = $N(N_p+1)$ ($N$: number of samples, $N_p$: number of parameters)
as discussed earlier in~\ref{sub:dgsm}. In the case of PCE-based computation of $\mathcal{T}(\theta_i)$,
$\mathcal{M}$ is the sum total of the number of training points used for constructing the sparse-basis
PCE and the number of evaluations used for its verification, denoted by $N_v$. In Table~\ref{tab:cost},
we provide a comparison of $\mathcal{M}$ for the two approaches in the case of 19 and 33 uncertain
parameters for the H$_2$/O$_2$ reaction kinetics application. 
%
\begin{table}[htbp]
\renewcommand{\arraystretch}{1.5}
\caption{A comparison of computational cost for the DGSM-based and Sobol'-based parametric 
sensitivity analysis in the case of 19 and 33 uncertain rate-controlling parameters.}
\begin{center}
\begin{tabular}{c|c|c|}
\cline{2-3}
& \multicolumn{2}{ c| }{$\mathbf{\#}$ \textbf{of Function Evaluations/Model Runs}~($\mathcal{M}$)} \\ \cline{2-3} 
& \textbf{19D} & \textbf{33D} \\  \hhline{===}
\multicolumn{1}{ |l|| }{\textbf{DGSM-based}, $\widehat{\mathcal{C}_i\mu_i}$} & 5(19+1) = 100 & 5(33+1) = 170 \\ 
\cline{1-3}
\multicolumn{1}{ |l|| }{\textbf{Total-effect Sobol'}, $\mathcal{T}(\theta_i)$} & 20 + $N_v$ & 
500 + $N_v$ \\ \cline{1-3}  
\end{tabular}
\end{center}
\label{tab:cost}
\end{table}
%
For the 19D case, estimating the converged estimates of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$)
required 5 samples. However, since finite difference was used in this work for estimating the gradient of the
model output, a total of 100 model runs were required. However, as mentioned earlier in~\ref{sub:dgsm}, depending
upon the application, it is possible to explore efficient techniques for gradient estimation such as those involving 
adjoint computation and automatic differentiation. Hence, there is scope for reduction in the computational cost
associated with the DGSM-based approach. Constructing the sparse-basis PCE requires only 20 samples in
the 19D parameter space with uncertain pre-exponents. However, additional model runs typically ranging from
$\mathcal{O}(10^{2})$ - $\mathcal{O}(10^{3})$ are needed to build a cross validation test suite to verify the 
accuracy of the PCE. The DGSM-based approach could thus yield computational gains especially when 
efficient gradient computation techniques are used. The comparison is significantly more favorable for the 
proposed DGSM-based framework in the higher dimensional case involving 33 uncertain parameters. 
Once again, converged estimates of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$) are obtained
using only 5 samples in the 33D parameter space. Hence, a total of 170 model runs are needed in this case
including for gradient computation using finite difference. Whereas, as discussed earlier,
constructing the sparse PCE itself requires model runs at 500 training points to sufficiently converge. Additional
model runs ($\mathcal{O}(10^{2})$ - $\mathcal{O}(10^{3})$) are required to further verify the accuracy of the 
resulting PCE. Note that the analysis was performed
for the fuel-rich case. However, similar trends are expected for the fuel-lean case and we do not include them here
for brevity.

In addition to the potential for computational cost reduction using the proposed framework, the comparative 
analysis also reveals the significant increase in computational effort associated with the PCE-based approach for
computing  the total-effect
Sobol' indices as the dimensionality increases from 19 to 33 despite using an efficient approach involving
sparse-basis PCEs. It can thus be understood that for higher dimensional problems involving hundreds of 
parameters, conventional approaches (even sparse-basis PCEs) can quickly become prohibitive and
the proposed DGSM-based approach could help
reduce the computational effort by several orders of magnitude.

>>>>

The computational cost is estimated in terms of the number of function evaluations or model runs,
denoted by $\mathcal{M}$ in each case. In the case of DGSM-based approach presented in this work, 
$\mathcal{M}$ = $N(N_p+1)$ ($N$: number of samples, $N_p$: number of parameters)
as discussed earlier in~\ref{sub:dgsm}. In the case of PCE-based computation of $\mathcal{T}(\theta_i)$,
$\mathcal{M}$ is the sum total of the number of training points used for constructing the sparse-basis
PCE and the number of evaluations used for its verification. In Table~\ref{tab:cost},
we provide a comparison of $\mathcal{M}$ for the two approaches in the case of 19 and 33 uncertain
parameters for the H$_2$/O$_2$ reaction kinetics application. 
%
\begin{table}[htbp]
\renewcommand{\arraystretch}{1.5}
\caption{A comparison of computational cost for the DGSM-based and Sobol'-based parametric 
sensitivity analysis in the case of 19 and 33 uncertain rate-controlling parameters.}
\begin{center}
\begin{tabular}{c|p{4cm}|p{4cm}|}
\cline{2-3}
& \multicolumn{2}{ c| }{$\mathbf{\#}$ \textbf{of Function Evaluations/Model Runs}~($\mathcal{M}$)} \\ \cline{2-3} 
& \centering \textbf{19D} & \hspace{18mm} \textbf{33D} \\  \hhline{===}
\multicolumn{1}{ |l|| }{\textbf{DGSM-based}, 
$\widehat{\mathcal{C}_i\mu_i}$} & \centering 100 + $N_{v_1}$ & \hspace{13mm} 170 + $N_{v_1}$ \\ 
\cline{1-3}
\multicolumn{1}{ |l|| }{\textbf{Total-effect Sobol'}, $\mathcal{T}(\theta_i)$} & \centering 20 + $N_{v_2}$ & 
\hspace{13mm} 500 + $N_{v_2}$ \\ \cline{1-3}  
\end{tabular}
\end{center}
\label{tab:cost}
\end{table}
%
For the 19D case, estimating the converged estimates of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$)
required 5 samples. However, since finite difference was used in this work for estimating the gradient of the
model output, a total of 100 model runs were required. Additionally, the screening procedure is continued for
one iteration using $N$ = 10 samples to ensure the convergence of $\widehat{\mathcal{C}_i\mu_i}$.
Since these additional runs are essentially used for verification, we denote them as $N_{v_1}$.
%As mentioned earlier in~\ref{sub:dgsm}, depending
%upon the application, it is possible to explore efficient techniques for gradient estimation such as those involving 
%adjoint computation and automatic differentiation. Hence, there is scope for a significant
%reduction in the computational cost
%associated with the DGSM-based approach. 
On the other hand, constructing the sparse-basis PCE requires only 20 samples in
the 19D parameter space with uncertain pre-exponents. Additional model runs ($N_{v_2}$)
 typically ranging from
$\mathcal{O}(10^{2})$ - $\mathcal{O}(10^{3})$ are needed to build a cross validation test suite to verify the 
accuracy of the PCE.  
The DGSM-based approach could thus yield computational gains especially when 
efficient gradient computation techniques are used. The comparison is significantly more favorable for the 
proposed DGSM-based framework in the higher dimensional case involving 33 uncertain parameters. 
Once again, converged estimates of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$) are obtained
using only 5 samples in the 33D parameter space. For verification, the screening procedure is
continued for one iteration by evaluating the model output at 10 samples in 33 dimensions. 
Therefore, a total of 340 model runs are needed in this case
including for gradient computation using finite difference as well as verification. Whereas, as discussed earlier,
constructing the sparse PCE itself requires model runs at 500 training points to sufficiently converge. Additional
model runs ($\mathcal{O}(10^{2})$ - $\mathcal{O}(10^{3})$) are required to further verify the accuracy of the 
resulting PCE. Note that the analysis was performed
for the fuel-rich case. However, similar trends are expected for the fuel-lean case and we do not include them here
for brevity.

Based on our findings, it appears that the proposed DGSM-based approach can offer a significant
computational advantage especially in higher dimensions due to multiple factors: (1) Computational
effort required for estimating $\mathcal{T}(\theta_i)$ is observed to increase substantially with dimensionality even
when using sparse-basis PCEs. Whereas, the DGSM-based estimates converge
with only a few samples ($N$ = 5) even for a relatively higher-dimensional case. Moreover, it must be noted that
the computational gains associated with the proposed DGSM-based approach could be enhanced
significantly by employing efficient gradient computation techniques such as those involving adjoints
and automatic differentiation as mentioned earlier in~\ref{sub:dgsm}; (2) The number of model runs
needed to verify the convergence of $\widehat{\mathcal{C}_i\mu_i}$ can be controlled
using the iterative strategy in this work
and is expected to be much smaller than the number of runs needed for verifying the accuracy of  
$\mathcal{T}(\theta_i)$ especially in high-dimensional settings. Specifically, for the 33D case,
model runs at 5 additional samples (corresponding to the first iteration) were used to verify the
convergence of $\widehat{\mathcal{C}_i\mu_i}$ as opposed to 1000 model evaluations used to verify
the accuracy of the PCE. 
It can thus be understood that for higher dimensional problems involving hundreds of 
parameters, conventional approaches (even sparse-basis PCEs) can quickly become prohibitive and
the proposed DGSM-based approach could help
reduce the computational effort by several orders of magnitude.

### 2018-10-28 Paper/disc.tex
..

Based on the results presented for the test problems and the kinetics
application, the proposed framework seems quite promising in its
potential for identifying the unimportant model inputs. \rebut{In fact, in 
the numerical tests and the chemical kinetics application presented in this
work, reasonable estimates of the screening metric are obtained during the
initial screening step with only 5 samples. This is indicative of a small
degree of
variability in the gradient of the model output with respect to individual
parameters, and therefore, a low variance of the Monte Carlo estimator.}
These observations could be exploited to construct efficient model surrogates
in a reduced input space. However, it is important to remain cognizant
about the limitations of the framework as well. 
For instance, the quantity of interest is required to be differentiable with
respect to each parameter in the considered domain. This condition once
satisfied, enhances the accuracy of the PCE-based surrogates as well.  
\rebut{Moreover, in the presence of severe non-linearity in the model output
leading to significant variability in its gradient in the considered input domain,
the number of samples and hence, the computational effort associated with
the estimation of screening metrics would expectedly 
increase. The extent of increase in the effort would of course depend upon the
application. In the case of rare events, the variance-based methods typically
lead to inaccurate estimates of the uncertainty in the model output. Since the
methodology proposed in this work is based on reducing computational
effort associated with variance-based sensitivity analysis, specifically, the
total-effect Sobol' index, it might lead to erroneous estimates of the
screening metrics and consequently the relative importance of the
uncertain parameters.}
Additionally, the proposed framework does not
account for the existence of possible correlations between the uncertain
inputs of the model. However, while the assumption of independent
inputs is not always justified, in many cases, correlations between
inputs are not well understood a priori, and assuming mutual independence
could be reasonable at least in initial screening using DGSMs. On the other
hand, if approximate correlations are known, we recommend using a Gaussian 
process or Kriging-based surrogate since it 
provides a means for incorporating the correlation between inputs.  
Implementation to applications involving strongly correlated parameters
could enhance the applicability of the proposed framework. 
We consider that to be a potential direction for future studies related
to this work.

>>>>

Based on the results presented for the test problems and the kinetics
application, the proposed framework seems quite promising in its
potential for identifying the unimportant model inputs. \rebut{In fact, in 
the numerical tests and the chemical kinetics application presented in this
work, reasonable estimates of the screening metric are obtained during the
initial screening step with a few samples. This is indicative of a small
degree of
variability in the gradient of the model output with respect to individual
parameters, and therefore, a low variance of the Monte Carlo estimator.}
These observations could be exploited to construct efficient model surrogates
in a reduced input space. However, it is important to remain cognizant
about the limitations of the framework as well. 
For instance, the quantity of interest is required to be differentiable with
respect to each parameter in the considered domain. This condition once
satisfied, enhances the accuracy of the PCE-based surrogates as well.  
\rebut{Moreover, in the presence of severe non-linearity in the model output
leading to significant variability in its gradient in the considered input domain,
the number of samples and hence, the computational effort associated with
the estimation of screening metrics would expectedly 
increase. The extent of increase in the effort would of course depend upon the
application. In the case of rare events, the variance-based methods typically
lead to inaccurate estimates of the uncertainty in the model output. Since the
methodology proposed in this work is based on reducing computational
effort associated with variance-based sensitivity analysis, specifically, the
total-effect Sobol' index, it might lead to erroneous estimates of the
screening metrics and consequently the relative importance of the
uncertain parameters.}
Additionally, the proposed framework does not
account for the existence of possible correlations between the uncertain
inputs of the model. However, while the assumption of independent
inputs is not always justified, in many cases, correlations between
inputs are not well understood a priori, and assuming mutual independence
could be reasonable at least in initial screening using DGSMs. On the other
hand, if approximate correlations are known, we recommend using a Gaussian 
process or Kriging-based surrogate since it 
provides a means for incorporating the correlation between inputs.  
Implementation to applications involving strongly correlated parameters
could enhance the applicability of the proposed framework. 
We consider that to be a potential direction for future studies related
to this work.

### 2018-10-28 Paper/examples.tex
..

\begin{figure}[htbp]
 \begin{center}
 \begin{tabular}{cc}
  \includegraphics[width=0.45\textwidth]{./Figures/err_samples_borehole}
  &
  \includegraphics[width=0.45\textwidth]{./Figures/errL2_samples_borehole}
  \\ (a) & (b)
  \end{tabular}
  \includegraphics[width=0.45\textwidth]{./Figures/pdf_comp_borehole}
  \\ (c)
\caption{(a) A comparison of order of the leave-one-out-error 
($\epsilon_{\mbox{\tiny{LOO}}}$) as a function of number of regression samples
used for constructing the PCE in 4, 5, and 7 dimensions. \rebut{The degree of
the PCE constructed using 50 training points was found to be 
3, 2,and 2 in the 4D, 5D, and 7D case respectively;
(b) Semilog plot of the relative L-2 error norm ($\epsilon_{\mbox{\tiny{L-2}}}$) 
for the PCEs constructed in 4D, 5D, and 7D};
and (c) A comparison of
PDFs of the discharge, $\mathcal{Q}$, generated using 10$^{6}$ samples from
the marginal distributions of the uncertain parameters in each case.} 
\label{fig:conv_bore}
\end{center}
\end{figure}
%
As expected, it is observed that the PCE constructed in the 4D space converges
at a much faster rate. For instance, if a PCE with $\mathcal{O}(10^{-4})$
accuracy is sought, we need function evaluations at only about 50 sample points
in the 4D parameter space whereas the number of samples needed in the full 7D
space seems much higher. Latin hypercube sampling (LHS) was used in
each case. It must be pointed out that the error in Figure~\ref{fig:conv_bore}(a)
is not expected to decrease monotonically with the increase in sample size
owing to the penalty term in the regularized optimization problem in Eq.~\ref{eq:reg}. 
\rebut{However, the plot of $\epsilon_{\mbox{\tiny{L-2}}}$ as a function of sample
size for the three surrogate models in Figure~\ref{fig:conv_bore}(b)
clearly demonstrates the trade-off between
computational effort associated with surrogate construction and its predictive 
accuracy. For a small sample size of 20, the relative  error in the case of 4D
PCE is found to be the smallest. However, as the sample size increases, the higher
order PCE becomes more accurate. Note that the same set of 50 samples in the
validation test suite were used in each case. Hence, when limited computational resources
are available, the proposed methodology could yield substantial gains due to fast
convergence as well as higher accuracy of the surrogate in the reduced space.
The gains are expected to increase exponentially as the dimensionality of the 
input space increases.}

@@

predictability. However, as the sample size increases, the full space is observed to

>>>>

\begin{figure}[htbp]
 \begin{center}
 \begin{tabular}{cc}
  \includegraphics[width=0.45\textwidth]{./Figures/err_samples_borehole}
  &
  \includegraphics[width=0.45\textwidth]{./Figures/errL2_samples_borehole}
  \\ (a) & (b)
  \end{tabular}
  \includegraphics[width=0.45\textwidth]{./Figures/pdf_comp_borehole}
  \\ (c)
\caption{(a) A comparison of order of the leave-one-out-error 
($\epsilon_{\mbox{\tiny{LOO}}}$) as a function of number of regression samples
used for constructing the PCE in 4, 5, and 7 dimensions. \rebut{The degree of
the PCE constructed using 50 training points was found to be 
3, 2,and 2 in the 4D, 5D, and 7D case respectively;
(b) Semilog plot of the relative L-2 error norm ($\epsilon_{\mbox{\tiny{L-2}}}$) 
for the PCEs constructed in 4D, 5D, and 7D};
and (c) A comparison of
PDFs of the discharge, $\mathcal{Q}$, generated using 10$^{6}$ samples from
the marginal distributions of the uncertain parameters in each case.} 
\label{fig:conv_bore}
\end{center}
\end{figure}
%
As expected, it is observed that the PCE constructed in the 4D space converges
at a much faster rate. For instance, if a PCE with $\mathcal{O}(10^{-4})$
accuracy is sought, we need function evaluations at only about 50 sample points
in the 4D parameter space whereas the number of samples needed in the full 7D
space seems much higher. Latin hypercube sampling (LHS) was used in
each case. It must be pointed out that the error in Figure~\ref{fig:conv_bore}(a)
is not expected to decrease monotonically with the increase in sample size
owing to the penalty term in the regularized optimization problem in Eq.~\ref{eq:reg}. 
\rebut{However, the plot of $\epsilon_{\mbox{\tiny{L-2}}}$ as a function of sample
size for the three surrogate models in Figure~\ref{fig:conv_bore}(b)
clearly demonstrates the trade-off between
computational effort associated with surrogate construction and its predictive 
accuracy. For a small sample size of 20, the relative  error in the case of 4D
PCE is found to be the smallest. However, as the sample size increases, the higher
order PCE becomes more accurate. This is expected since the higher order PCE
inherently contains more information pertaining to the input-output relationship.
Note that the same set of 50 samples in the
validation test suite were used in each case. Hence, when limited computational resources
are available, the proposed methodology could yield substantial gains due to fast
convergence as well as higher accuracy of the surrogate in the reduced space.
The gains are expected to increase significantly as the dimensionality of the 
input space increases.}

@@

predictability. However, as the sample size increases, the FSS is observed to
